---
title: "Distinguish 6 clones"
author: "Ruben Hermann"
date: '2022-04-08'
output: html_document
---

Loading packages and emptying list

```{r, include=FALSE}
rm(list=ls())

library(keras)
library(tidyverse)
```

Loading in the data

```{r}
#Reading in the model configuration adta
name <-paste('../Data/Model configuration',sep="")
  setwd(name)
  temp= list.files(pattern=".csv")
  listofdf<-list()
  for (i in 1:length(temp)) {
    assign(temp[i],read.csv(temp[i],))
    listofdf[[i]]=read.csv(temp[i])}
  names(listofdf) <- temp

#Binding the data frames of the different clonal lineages together
all_test<-rbind(CR1_data_for_test.csv,CR2_data_for_test.csv,CR3_data_for_test.csv,CR4_data_for_test.csv,CR6_data_for_test.csv,CR7_data_for_test.csv)  
all_train<-rbind(CR1_data_for_training.csv,CR2_data_for_training.csv,CR3_data_for_training.csv,CR4_data_for_training.csv,CR6_data_for_training.csv,CR7_data_for_training.csv)  
all_val<-rbind(CR1_data_for_validation.csv,CR2_data_for_validation.csv,CR3_data_for_validation.csv,CR4_data_for_validation.csv,CR6_data_for_validation.csv,CR7_data_for_validation.csv) 

#deleting the X column
all_test <- select(all_test,-X)
all_train <- select(all_train,-X)
all_val <- select(all_val,-X)

#Randomizing the order of samples in the data frames
all_train <- all_train[sample(nrow(all_train)),]
all_val <- all_val[sample(nrow(all_val)),]
all_test <- all_test[sample(nrow(all_test)),]
```

Scaling the data according to the mean and sd of the training data

```{r}
#First I need to normalize the training data
 ID <- all_train$ID
 all_train <- select(all_train,-ID)
 mean <- apply(all_train,2,mean)
 std <- apply(all_train,2,sd)
 all_train_sc <- scale(all_train,center=mean,scale=std)
 all_train_sc <- as.data.frame(all_train_sc)
 all_train_sc$ID <- ID
 all_train$ID <- ID

 
#Now I want to delete some features that do not help distinguishing the object
 list=names(all_train_sc)
 list <- list[list != 'ID'] # remove 'ID' which tells identity of the isolates
 list <- list[list != 'Object.Number'] # removing the object number it is not needed
 list <- list[!str_detect(list, 'Saturation')] #remove Saturation Perc as it always is 0
 list <- list[!str_detect(list, 'Bkgd')] #remove Background information
 #apply the list back to the data frame
 all_train_sc <- all_train_sc %>% dplyr::select(all_of(list),ID)
 
########################################################################################
 #######################################################################################
 #######################################################################################
 
 #Now I need to normalize the other data frames as well
 #First the validation
 ID <- all_val$ID
 all_val <- select(all_val,-ID)
 all_val_sc <- scale(all_val,center=mean,scale=std)
 all_val_sc <- as.data.frame(all_val_sc)
 all_val$ID <- ID
 all_val_sc$ID <- ID
 
 #Now I want to delete some features that do not help distinguishing the object
 all_val_sc <- all_val_sc %>% dplyr::select(all_of(list),ID)
 
 
 ########################################################################################
 #######################################################################################
 #######################################################################################
 #Second the test
 ID <- all_test$ID
 all_test <- select(all_test,-ID)
 all_test_sc <- scale(all_test,center=mean,scale=std)
 all_test_sc <- as.data.frame(all_test_sc)
 all_test$ID <- ID
 all_test_sc$ID <- ID

 #Now I want to delete some features that do not help distinguishing the object
 all_test_sc <- all_test_sc %>% dplyr::select(all_of(list),ID)
 
```

Setting the data frames to the right format using keras commands

```{r}
no_ID <- length(unique(all_train_sc$ID))
no_vars <- length(list)

 #Formating the training and validation data
 train_x <- all_train_sc[,1:(no_vars)] %>% as.matrix()
 train_y <- all_train_sc$ID
 train_y <- to_categorical(train_y, no_ID)
 
 val_x <- all_val_sc[,1:(no_vars)] %>% as.matrix()
 val_y <- all_val_sc$ID
 val_y <- to_categorical(val_y, no_ID)
 
 test_x <- all_test_sc[,1:(no_vars)] %>% as.matrix()
 test_y <- all_test_sc$ID
 test_y <- to_categorical(test_y, no_ID)  
```

Creating the 6clone model

```{r}

model <- keras_model_sequential() 
model %>% 
     layer_dense(units = 120,activation = 'relu', input_shape = c(no_vars)) %>% 
     layer_dropout(rate = 0.5) %>% 
     layer_dense(units = 30, activation = 'relu') %>%
     layer_dropout(rate = 0.5) %>%  
     layer_dense(units = no_ID, activation = 'softmax')
   

   
model %>% compile(
     loss = 'categorical_crossentropy',
     optimizer = optimizer_rmsprop(learning_rate = 0.001),
     metrics = 'accuracy'
   )
   
   
   
   
hist_<-model %>% fit(
     train_x, 
     train_y, 
     epochs = 70, 
     batch_size = 800, 
     validation_data = list(val_x, val_y ),
     #verbose = 0,
   )

   # ---------------accuracy of the model -------------------#
model %>% evaluate(test_x, test_y,verbose=0)

predict <- model %>% predict(test_x)

save_model_hdf5(model,"Distinguish 6 clones with feature values.h5")
```

Checking the recall rate of the predictions
First a general check on the recall rate predicted clonal lineage and on the recall rate of the phenotype of the clonal lineage

Second, using a higher certainty (90%) for the model when predicting a clonal lineage and checking the recall rate

```{r}
pred<-data.frame(real_ID=(all_test_sc$ID+1),pred_ID=6,acc=0,recall=0,real_phe=0,pred_phe=0,recall_phe=0)


for (i in 1:nrow(predict)){
  pred$pred_ID[i]=which.max(predict[i,])
  pred$acc[i]=predict[i,which.max(predict[i,])]
  pred$recall[i]<- pred$pred_ID[i]==pred$real_ID[i]
      if (pred$real_ID[i]<4) {pred$real_phe[i]="def"} 
      if (pred$real_ID[i]==4) {pred$real_phe[i]="udef"}
      if (pred$real_ID[i]>4) {pred$real_phe[i]="inter"}
      if (pred$pred_ID[i]<4) {pred$pred_phe[i]="def"} 
      if (pred$pred_ID[i]==4) {pred$pred_phe[i]="udef"} 
      if (pred$pred_ID[i]>4) {pred$pred_phe[i]="inter"} 
  pred$recall_phe[i] <- pred$real_phe[i]==pred$pred_phe[i]
 }
pred

sum(pred$recall)/nrow(pred)
sum(pred$recall_phe/nrow(pred))
mean(pred$acc)

#Creating a new data frame
pred_high <- data.frame(real_ID=0,pred_ID=0,acc=0,recall=0,real_phe=0,pred_phe=0,recall_phe=0)

#taking only the prediction with a cartainty of 90%
for (i in 1:nrow(pred)) {
  if (pred$acc[i]>0.9)
  { pred_high[i,]=pred[i,]}
}

pred_high<-na.omit(pred_high)
nrow(pred_high)/nrow(pred)
sum(pred_high$recall)/nrow(pred_high)
sum(pred_high$recall_phe/nrow(pred_high))
```